WSClean: an implementation of a fast, generic wide-field imager

### 2014-02-02 articles/2014/Article-WSClean/wsclean.tex
Work on article

New generation of wide-field observatories are producing data sets that are orders of magnitude larger than before. Examples of such telescopes include the Murchison Widefield Array (MWA, \citealt{mwa}), the upgraded Jansky Very Large Array (JVLA) and the Low-Frequency Array (LOFAR, \citealt{lofar-2013}). At the MWA, some of the data is reduced with the w-projection algorithm implementation of the Common Astronomy Software Applications (CASA, \citealt{casa}). This is an attractive option because of its many features, high quality and free availability. However, at the MWA we have seen that a 2-min snapshot observation at low-elevation can take up to tens of hours with CASA's w-projection algorithm. Another option for reducing MWA data is the Real-Time System (RTS), which has been designed as an efficient calibrating and imaging pipeline specifically for MWA data \citep{rts-mwa}. It can use GPUs for improving efficiency, and uses a w-projection kernel during imaging to correct for the w-term. The kernel can be kept small, because the data is processed in snapshots that have been phase-centered to Zenith. The RTS can not perform deconvolution as of yet.

@@

\subsection{Discretization of w} \label{sec:gridding-w}
\begin{figure}
\begin{center}
\includegraphics[width=8cm]{img/aliasing-example-annotated}
\caption{Aliasing artefacts caused by insufficient $w$-layers. WSClean was set to use 12 $w$-layers. The centre of the image is at 10\degree zenith angle. This would normally require imaging with $\sim195$ $w$-layers. Original sources are 1~Jy, ghost sources are approximately 20\%. Each source produces two ghost sources, but because they re-appear after a major cleaning cycle, they are eventually cleaned and produce ghost sources by themselves. }
\label{fig:aliasing-example}
\end{center}
\end{figure}
While the discretization of $u$ and $v$ is similar to conventional imaging, the discretization of $w$ defines the number of w-layers that need to be processed. For this, one can use a rule similar to \eqref{eq:when-2d-is-valid}, and make sure that the phase difference for two subsequent discretized $w$-values, $w_A$ and $w_B$,  is less than one radian. This results in the constraint
\begin{equation} \label{eq:minimum-w-distance}
\left|\left(w_A - w_B\right) 2\pi (\sqrt{1-l^2-m^2}-1)\right| \ll 1.
\end{equation}
This suggest that a uniform discretization in $w$ is optimal TODO in contrast to some article? -- could be explained by having more samples at smaller $w$s. From Eq.~\eqref{eq:minimum-w-distance}, the required number of layers can be derived and is given by
\begin{equation} \label{eq:nwlayers-bound}
 N_\textrm{wlayers} \gg 2\pi \left(w_e - w_0\right) \max_{l,m} \left(1 - \sqrt{1-l^2-m^2}\right).
\end{equation}
Actual values for the right-hand side can differ a lot. The value of $w_e - w_0$ is influenced by the coplanarity of the array, the zenith angle and the wavelength, while the value of the $\max_{l,m}$ term is influenced by the angular size of the image. For the MWA, a typical value of $w_e - w_0$ is $\sim 10$ at zenith, but is already $\sim 400$ at a zenith angle of $30^{\circ}$. [TODO table of arrays with $w$ values at zenith/more angles/more freq?] For a typical full-field-of-view image of a MWA observation of $3072\times 3072$ pixels of $0.75''$ size, the $\max_{l,m}$ term is $0.68$. This implies that, in the case of the MWA, tens of $w$-layers are required at zenith and hundreds at lower elevations. The number of $w$-layers has a large effect on the performance of the $w$-stacking algorithm, and will be discussed further in the next sections.

@@

\subsection{Time complexity of w-stacking}
It is useful to analyse the time complexity of w-stacking and compare it with w-projection, to understand which algorithm performs better in a given situation. We will use the following symbols: $N_\textrm{wlay}$ is the number of w-layers for w-stacking, $N_\textrm{pix}$ is the number of pixels in the image along each dimension, $N_\textrm{vis}$ is the number of visibilities and $N_\textrm{kern}$ is the size of the antialiasing kernel (see \S\ref{sec:gridding}), $N_\textrm{wkern}$ is the size of the w-kernel for w-projection, $w_e$ is the maximum $w$ value and $\alpha_\textrm{FOV}$ is the imaging field of view.

@@

\section{The WSClean imager implementation} \label{sec:implementation}
With the purpose of testing new algorithms for the imaging of MWA data, we have written a new imager around the $w$-stacking algorithm. Nevertheless, the imager is not specialized for the MWA, and has been succesfully used for other telescopes (TODO menton more).

The new imager, called ``WSClean'', is written in the C++ language. It reads visibilities from CASA measurement sets and writes output images to Flexible Image Transport System (FITS) files. Several steps are multi-threaded: reading and writing; gridding from and to different $w$-layers; performing the Fast Fourier transforms; and H\"ogbom Clean iterations (peak-finding and image subtraction, \citealt{hogbom-clean}). For the latter, intrinsics are used as well. Because of these optimizations, minor Cleaning iterations are performed at a rate of hundreds per second at $3072\times3072$. This is fast enough to make the Clark Clean optimization \citep{clark-clean}, which consists of considering only a subset of pixels with a reduced PSF, less relevant. However, the PSF varies for imaging with non-zero $w$-values, and subtracting a constant PSF in image space leads to inaccuracies. After a number of minor iterations, it is therefore beneficial to invert the model back to the visibilities via prediction, and subtract the model directly from the visibilities \citep{wprojection-cornwell}. This is similar to the Cotton-Schwab cleaning method \citep{cotton-schwab-clean}.

@@

\subsection{W-snapshots} \label{sec:snapshot-imaging}
\begin{figure*}
\begin{center}
\includegraphics[width=8cm]{img/vela-normal-projection}
\includegraphics[width=8cm]{img/vela-zenith-projection}
\caption{13-min MWA observation of supernova remnants Vela and Puppis A. Left: normal projection centred on Puppis A, right: phase-rotated to zenith to reduce $w$-values, recentred on Puppis A during imaging. Both Stokes-I images have been made with WSClean using the Cotton-Schwab clean algorithm. No beam correction was applied. TODO time difference and more discussion, replace with deeper images.}
\label{fig:vela-projection-example}
\end{center}
\end{figure*}
As shown, all-sky imaging can be quite efficient with snapshot imaging towards zenith. However, if one is not interested in imaging the whole sky, and the direction of interest is far from zenith, the computational overhead of making all-sky images is undesirable.

@@

In all analyses CASA stable release 42.0, revision 26465 was used, which was released September 2013. It includes performance improvements to the \texttt{clean} task which have recently been made. The latest version as of February 2014 of WSClean was used. The tests were run on a high-end desktop, with a 3.20-GHz Intel Core i7-3930K processor and 32 GB of memory. The data are stored on a multi-disk array with 5 spinning hard disks, which has a combined read rate of about 450 MB/s. 
\begin{table}%
\caption{Parameter values used during benchmarks, unless otherwise mentioned.} \label{tbl:default-parameters}%
\begin{center}\begin{tabular}{rl}%
\hline
Array & MWA 128T \\
Image size & $3072 \times 3072$ \\
Angular resolution & $0.72'$ \\
Number of visibilities & $3.5 \times 10^8$ \\
Time resolution & 2~s \\
Frequency resolution & 40~kHz \\
Observation duration & 112 s\\
Bandwidth & 30.72 MHz (768 channels)\\
Central frequency & 182 MHz \\
Zenith angle at phase centre & 10\degree \\
Max $w$-value for phase centre & 172 $\lambda$ (283 m) \\
Nr. polarizations in set & 4 \\
Imaged polarization & XX \\
Imaging mode & Multi-frequency synthesis \\
Weighting & uniform \\
Data size & 18 GB \\
\hline
\end{tabular}\end{center}\end{table}

@@

We performed measurements to assess the accuracy of WSClean and CASA. To do so, we have simulated a MWA observation without system noise, but with hundred sources of 1~Jy in a 20\degree diameter area. A unitary primary beam is assumed. We imaged the simulated set both with WSClean and CASA, using Cotton-Schwab cleaning to a threshold of 10~mJy. The two imagers calculate slightly different restoring (synthesized) beams, hence to avoid bias the restoring beams are fixed. Other imaging parameters are given in Table~\ref{tbl:default-parameters}. The Aegean program \citep{aegean-hancock-2012} was used to perform source detection on the produced images. Sidelobe noise of the residual 10~mJy noise triggers a few false detections. These are ignored.

@@

In the $ZA=10\degree$ case, CASA is slightly less accurate. This is consistent, and can actually be observed in the images as shown in Fig.~\ref{fig:residuals}. TODO this might be caused by a bug; check: We have also made an image with the technique of recentring a zenith phase-centred visibility set, as described in Sect.~\ref{sec:snapshot-imaging}. Its accuracy is lower compared to the normal projection. This is due to the fact that the sources are further away from the phase centre, and sources have therefore on average higher $(l,m)$-values, thereby increasing the average $w$-correction. Nevertheless, the reprojection technique needs on average fewer $w$-layers to reach the same level of accuracy.

@@

Many parameters affect the accuracy more than the observed differences between the methods. These include the number of major iterations, $w$-planes/layers, restoring beam, image size, weighting, kernel size and grid oversampling. Therefore, the differences are not very significant. The results do show that synthesis imaging using common imaging parameters results in an approximate flux-density standard error of at least $\sim1\%$.

@@

Fig.~\ref{fig:timing-nsamples} shows the wall-clock time to image a 112-s MWA snapshot as a function of the size of the visibility set. Results for imaging at 10\degree zenith angle (ZA) and at zenith are shown. The size of the visibility set was varied by changing the time resolution. This affects the number of visibilities to be gridded without changing the maximum $w$-value. For larger sets, both methods show a linear time dependency on the number of visibilities. This implies that gridding or reading dominates the cost. In that situation, the $w$-stacking implementation is 7.9 times faster at $\textrm{ZA}=10\degree$ and 2.9 times faster at $\textrm{ZA}=0\degree$. At $\textrm{ZA}=10$\degree, the performance of WSClean is approximately constant for sets with less than about $10^8$ visibilities, which corresponds with a single polarization of 16~s of MWA data in multi-frequency synthesis mode. At that point, the FFTs are the dominant cost in the algorithm. At zenith, the two methods are expected to perform almost identical. The factor of 2-3 difference in Fig.~\ref{fig:timing-nsamples} could be due to different choices in optimizations.

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{img/benchmark-zenith-angle/za}
\caption{Effect of zenith angle (ZA) on imaging performance. Image width and height are $2048$ for the dashed line and $3072$ for the solid line. CASA fails to image observations with ZA $\ge 15$ with a $3072$ image size. ZA dramatically affects the computational cost of w-projection. }
\label{fig:timing-zenith-angle}
\end{center}
\end{figure}

In Fig.~\ref{fig:timing-zenith-angle}, the computational cost as function of the zenith angle is plotted.

@@

@@

\subsection{Gridding} \label{sec:gridding}
During gridding of the visibilities in $uv$-space, it is useful to apply a gridding convolution kernel. A common kernel function is a windowed sinc function. A low-pass filter decreases the flux of sources outside the field of view, and thus helps to attenuate aliased ghost sources and sidelobes (see \citealt{post-correlation-filtering}) that arise because of the assumed periodicity during the inverse FFT. By supersampling, a convolution kernel also makes it possible to place samples more accurately at their $uv$ position, thereby lowering decorrelation.

The prolate spheroidal wave function (PSWF) is generally considered to be the optimal windowing function for gridding (TODO cite). CASA convolves samples with a PSWF of 7 pixels total width during gridding. When using a variable kernel size, a PSWF is quite complicated and computational expensive to calculate. WSClean currently uses a Kaiser-Bessel (KB) window function, which is easy and fast to compute, and its properties are similar to a PSWF.

@@

The type of window function has no effect on the gridding performance, but gridding has a quadratic time dependency on the size of the kernel. Fig.~\ref{fig:timing-kernelsize} shows that this quadratic dependency becomes apparent for kernel sizes $\ge 15$. Decreasing the kernel size to values below $7$ has little effect, because reading the samples takes more time than gridding with small kernels. We have not noticed much benefit of larger kernels except in rare cases where a bright source lies just outside the imaged field of view. Therefore, WSClean uses a default of 7 pixels for the gridding kernel. It can be increased when necessary.

@@

>>>>

New generation of wide-field observatories are producing data sets that are orders of magnitude larger than before. Examples of such telescopes include the Murchison Widefield Array (MWA, \citealt{mwa}), the upgraded Jansky Very Large Array (JVLA) and the Low-Frequency Array (LOFAR, \citealt{lofar-2013}). At the MWA, some of the data is reduced with the w-projection algorithm implementation of the Common Astronomy Software Applications (CASA, \citealt{casa}). This is an attractive option because of its many features, high quality and free availability. However, at the MWA we have seen that imaging a 2-min snapshot observation at low-elevation can take up to tens of hours with CASA's w-projection algorithm. Another option for reducing MWA data is the Real-Time System (RTS), which has been designed as an efficient calibrating and imaging pipeline specifically for MWA data \citep{rts-mwa}. It can use GPUs for improving efficiency, and uses a w-projection kernel during imaging to correct for the w-term. The kernel can be kept small, because the data is processed in snapshots that have been phase-centered to Zenith. The RTS can not perform deconvolution as of yet.

@@

\subsection{Discretization of w} \label{sec:gridding-w}
\begin{figure}
\begin{center}
\includegraphics[width=8cm]{img/aliasing-example-annotated}
\caption{Aliasing artefacts caused by insufficient $w$-layers in a simulated field. WSClean was set to use 12 $w$-layers. The centre of the image is at 10\degree zenith angle, which would normally require $\sim195$ $w$-layers. Original sources are 1~Jy, ghost sources are approximately 20\%. Each source produces two ghost sources, but because they re-appear after a major cleaning cycle, they are eventually cleaned and produce more ghost sources. }
\label{fig:aliasing-example}
\end{center}
\end{figure}
While the discretization of $u$ and $v$ is similar to conventional imaging, the discretization of $w$ defines the number of w-layers that need to be processed. For this, one can use a rule similar to \eqref{eq:when-2d-is-valid}, and make sure that the phase difference for two subsequent discretized $w$-values, $w_A$ and $w_B$,  is less than one radian. This results in the constraint
\begin{equation} \label{eq:minimum-w-distance}
\left|\left(w_A - w_B\right) 2\pi (\sqrt{1-l^2-m^2}-1)\right| \ll 1.
\end{equation}
This suggest that a uniform discretization in $w$ is optimal TODO in contrast to some article? -- could be explained by having more samples at smaller $w$s. From Eq.~\eqref{eq:minimum-w-distance}, the required number of layers can be derived and is given by
\begin{equation} \label{eq:nwlayers-bound}
 N_\textrm{wlayers} \gg 2\pi \left(w_e - w_0\right) \max_{l,m} \left(1 - \sqrt{1-l^2-m^2}\right).
\end{equation}
Actual values for the right-hand side can differ a lot. The value of $w_e - w_0$ is influenced by the coplanarity of the array, the zenith angle and the wavelength, while the value of the $\max_{l,m}$ term is influenced by the angular size of the image. For the MWA, a typical value of $w_e - w_0$ is $\sim 10$ at zenith, but is already $\sim 400$ at a zenith angle of $30^{\circ}$. [TODO table of arrays with $w$ values at zenith/more angles/more freq?] For a typical full-field-of-view image of a MWA observation of $3072\times 3072$ pixels of $0.75''$ size, the $\max_{l,m}$ term is $0.68$. This implies that, in the case of the MWA, tens of $w$-layers are required at zenith and hundreds at lower elevations. The number of $w$-layers has a large effect on the performance of the $w$-stacking algorithm, and will be discussed further in the next sections.

@@

\subsection{Time complexity of w-stacking} \label{sec:time-complexity}
It is useful to analyse the time complexity of w-stacking and compare it with w-projection, to understand which algorithm performs better in a given situation. We will use the following symbols: $N_\textrm{wlay}$ is the number of w-layers for w-stacking, $N_\textrm{pix}$ is the number of pixels in the image along each dimension, $N_\textrm{vis}$ is the number of visibilities and $N_\textrm{kern}$ is the size of the antialiasing kernel (see \S\ref{sec:gridding}), $N_\textrm{wkern}$ is the size of the w-kernel for w-projection, $w_e$ is the maximum $w$ value and $\alpha_\textrm{FOV}$ is the imaging field of view.

@@

\section{The WSClean imager implementation} \label{sec:implementation}
With the purpose of testing new algorithms for the imaging of MWA data, we have written a new imager around the $w$-stacking algorithm. The imager is not specialized for the MWA, and has been succesfully used for other telescopes (TODO menton more).

The new imager, called ``WSClean'', is written in the C++ language. It reads visibilities from CASA measurement sets and writes output images to Flexible Image Transport System (FITS) files. Several steps are multi-threaded: reading and writing; gridding from and to different $w$-layers; performing the Fast Fourier transforms; and H\"ogbom Clean iterations (peak-finding and image subtraction, \citealt{hogbom-clean}). For the latter, intrinsics are used as well. Because of these optimizations, minor Cleaning iterations with an image size of $3072\times3072$ are performed at a rate of hundreds per second. This is fast enough to make the Clark Clean optimization \citep{clark-clean}, which consists of considering only a subset of pixels with a reduced PSF, less relevant. However, the PSF varies for imaging with non-zero $w$-values, and subtracting a constant PSF in image space leads to inaccuracies. After a number of minor iterations, it is therefore beneficial to invert the model back to the visibilities via prediction, and subtract the model directly from the visibilities \citep{wprojection-cornwell}. This is similar to the Cotton-Schwab cleaning method \citep{cotton-schwab-clean}.

@@

\subsection{W-snapshots} \label{sec:snapshot-imaging}
\begin{figure*}
\begin{center}
\includegraphics[width=8cm]{img/vela-normal-projection}
\includegraphics[width=8cm]{img/vela-zenith-projection}
\caption{13-min MWA observation of supernova remnants Vela and Puppis A. Left: normal projection centred on Puppis A, right: phase-rotated to zenith to reduce $w$-values, recentred on Puppis A during imaging. Both Stokes-I images have been made with WSClean using the Cotton-Schwab clean algorithm. No beam correction was applied. TODO time difference and more discussion, replace with deeper images.}
\label{fig:vela-projection-example}
\end{center}
\end{figure*}
As discussed, all-sky snapshot imaging with zenith as phase centre is quite efficient. However, if one is not interested in imaging the whole sky, and the direction of interest is far from zenith, the computational overhead of making all-sky images is undesirable.

@@

The used software version of CASA is ``stable release 42.0, revision 26465'', which was released September 2013. It includes performance improvements to the \texttt{clean} task which have recently been made. For WSClean, the latest version as of February 2014 was used. The tests were run on a high-end desktop, with a 3.20-GHz Intel Core i7-3930K processor with 6 cores and 32 GB of memory. The data are stored on a multi-disk array with 5 spinning hard disks, which has a combined read rate of about 450 MB/s. 
\begin{table}%
\caption{Parameter values used during benchmarks, unless otherwise mentioned.} \label{tbl:default-parameters}%
\begin{center}\begin{tabular}{rl}%
\hline
Array & MWA 128T \\
Image size & $3072 \times 3072$ \\
Angular resolution & $0.72'$ \\
Number of visibilities & $3.5 \times 10^8$ \\
Time resolution & 2~s \\
Frequency resolution & 40~kHz \\
Observation duration & 112 s\\
Bandwidth & 30.72 MHz (768 channels)\\
Central frequency & 182 MHz \\
Zenith angle at phase centre & 10\degree \\
Max $w$-value for phase centre & 172 $\lambda$ (283 m) \\
Nr. polarizations in set & 4 \\
Imaged polarization & XX \\
Imaging mode & Multi-frequency synthesis \\
Weighting & uniform \\
Data size & 18 GB \\
\hline
\end{tabular}\end{center}\end{table}

@@

To asses the accuracy of WSClean and CASA, we have simulated a MWA observation without system noise, but with hundred sources of 1~Jy in a 20\degree diameter area. A unitary primary beam is assumed. We imaged the simulated set both with WSClean and CASA, using Cotton-Schwab cleaning to a threshold of 10~mJy. The two imagers calculate slightly different restoring (synthesized) beams, hence to avoid bias the restoring beams are fixed. Other imaging parameters are given in Table~\ref{tbl:default-parameters}. The Aegean program \citep{aegean-hancock-2012} is used to perform source detection on the produced images. Sidelobe noise of the residual 10~mJy source structures triggers a few false detections. These are ignored.

@@

In the $\textrm{ZA}=10\degree$ case, CASA is slightly less accurate. We have noticed this consistently at higher zenith angles, and it can actually be observed in the produced images, as shown in Fig.~\ref{fig:residuals}. TODO the following might be caused by a bug; check: We have also made an image with the technique of recentring a zenith phase-centred visibility set, as described in Sect.~\ref{sec:snapshot-imaging}. Its accuracy is lower compared to the normal projection. This is due to the fact that the sources are further away from the phase centre, and sources have therefore on average higher $(l,m)$-values, thereby increasing the average $w$-correction. Nevertheless, the reprojection technique needs on average fewer $w$-layers to reach the same level of accuracy.

@@

Many parameters affect the accuracy more than the observed differences between the methods. These include the number of major iterations, $w$-planes/layers, restoring beam, image size, weighting, kernel size and grid oversampling. Therefore, the differences in accuracy are not very significant. The results do show that synthesis imaging using common imaging parameters results in an approximate flux-density standard error of at least $\sim1\%$.

@@

Fig.~\ref{fig:timing-nsamples} shows the wall-clock time to image a 112-s MWA snapshot as a function of the size of the visibility set. Results for imaging at 10\degree zenith angle (ZA) and at zenith are shown. The size of the visibility set was varied by changing the time resolution. This affects the number of visibilities to be gridded without changing the maximum $w$-value. For larger sets, both methods show a linear time dependency on the number of visibilities. This implies that gridding or reading dominates the cost. In that situation, the $w$-stacking implementation is 7.9 times faster at $\textrm{ZA}=10\degree$ and 2.6 times faster at $\textrm{ZA}=0\degree$. With small data volumes, the FFTs start to dominate the cost. This is visible from Fig.~\ref{fig:timing-nsamples} by the flattening towards the left. At that point, WSClean is in both cases approximately 3 times faster. At zenith, the two methods are expected to perform almost identical. The factor of 2-3 difference in Fig.~\ref{fig:timing-nsamples} could be due to different choices in optimizations.

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{img/benchmark-zenith-angle/za}
\caption{Effect of zenith angle (ZA) on imaging performance. Image pixel size is $2048$ for the dashed lines and $3072$ for the solid lines. CASA fails to image observations with ZA $\ge 15$ with a $3072$ pixel image size. }
\label{fig:timing-zenith-angle}
\end{center}
\end{figure}

In Fig.~\ref{fig:timing-zenith-angle}, the computational cost as function of zenith angle is plotted. It shows that, as expected from \S\ref{sec:time-complexity}, the $w$-projection algorithm is more affected by the increased $w$-values, leading to differences of more than an order of magnitude at zenith angles of $\gtrapprox 20$\degree. Additionally, at higher zenith angles, the $w$-kernels become too large to be able to make images of 3072 pixels or larger. Using the recentering technique to decrease the $w$-values makes the computational cost approximately constant. However, the additional time required to rephase and regrid the measurement set makes it only worthwhile for images $\ge3072$ pixels and $\textrm{ZA}\ge15\degree$. TODO: this is quite heavily depending on the speed of \texttt{chgcentre}, and that might still be improvable...

@@

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{img/benchmark-resolution/resolution}
\caption{...}
\label{fig:timing-resolution}
\end{center}
\end{figure}

@@

\subsection{Gridding} \label{sec:gridding}
A gridding convolution kernel improves the accuracy of gridding in $uv$-space. A common kernel function is a windowed sinc function, which acts as a low-pass filter. This decreases the flux of sources outside the field of view, and thus helps to attenuate aliased ghost sources and sidelobes that arise because of the assumed periodicity during the inverse FFT \citep{post-correlation-filtering}. By supersampling, a convolution kernel also makes it possible to place samples more accurately at their $uv$ position, thereby lowering decorrelation.

The prolate spheroidal wave function (PSWF) is generally considered to be the optimal windowing function for gridding \citep{fourier-kernel-selection-1991}. CASA convolves samples with a PSWF of 7 pixels total width during gridding. When using a variable kernel size, a PSWF is quite complicated and computational expensive to calculate. WSClean currently uses a Kaiser-Bessel (KB) window function, which is easy and fast to compute, and is a good approximation for the PSWF \citep{fourier-kernel-selection-1991}.

@@

The type of window function has no effect on the gridding performance, but gridding has a quadratic time dependency on the size of the kernel. Fig.~\ref{fig:timing-kernelsize} shows that this quadratic dependency becomes apparent for kernel sizes $\gtrapprox 15$. Decreasing the kernel size to values below $7$ has little effect on performance, because with small kernels reading the samples takes more time than gridding. We have not noticed much benefit of larger kernels except in rare cases where a bright source lies just outside the imaged field of view. Therefore, WSClean uses a default of 7 pixels for the gridding kernel. It can be increased when necessary.

@@

TODO: discuss at what is the optimal time to split measurement sets into when using recentering technique.
